Simulation Study Document

### 2021-08-21 Permanova-Real-Data-Section.Rmd
Adding files from R

---
title: 'Permanova: Real Data Analysis'
author: "Paul Harmon"
date: "7/13/2021"
output: html_document
---

@@

## NEED TO STANDARDIZE THE DATA FIRST (MAKE UNITLESS)
CCscale <- CC2015_filter %>% lapply(scale) %>% as_tibble()

out <- MultiPermanova(CCscale)

@@

  #Visualize the results and store in a list
cc <- CC2015 %>% filter(BASIC2015 %in% c(15,16,17)) 
  cc$NAME[which(df3$PValues < 0.05)]
  df3$Name <- cc$NAME
  sigplot2 <- ggplot(df3, aes(Holdout, Fstat, label = Name)) + geom_point(aes(color = Colors), size = 2) + geom_line(alpha = 0.5, color = "grey") + theme_bw() + ggtitle("CMD: P-Values From Adonis")
  sigplot2 %>% ggplotly()

## Scaled with Per-Capita Features

@@

```{r, cache = TRUE, eval = FALSE}
CC2015_filter_PC <- CC2015_filter %>% mutate(S.ER.DPC = S.ER.D/FACNUM, NONS.ER.DPC = NONS.ER.D/FACNUM, STAFFPC = PDNFRSTAFF/FACNUM)

@@

## Plots with Permutations

```{r}
library(tidyverse)
library(ggplot2)
library(ggridges)

histlist1 <- list()
histlist2 <- list()

for(j in 1:length(out$tsnelist)){
  histlist1[[j]] <- out$tsnelist[[j]]$f.perms %>% as_tibble() %>% ggplot(aes(V1)) + geom_histogram(bins = 10)
  histlist2[[j]] <-  out$cmdlist[[j]]$f.perms %>% as_tibble() %>% ggplot(aes(V1)) + geom_density() + scale_x_log10()
}

f1Fs_long %>% ggplot(mapping=aes(y=Number, x=value, fill = factor(stat(quantile)))) +
    stat_density_ridges(geom = "density_ridges_gradient",
    calc_ecdf = TRUE,
    quantiles = 0.95, 
    alpha=0.3,
    scale=2
    ) +
    scale_fill_manual(
    name = "Probability", values = c("#A0A0A0A0", "#FF0000A0"),
    labels = c("(0, 0.95]", "(0.95, 1]")
    )  +
   scale_x_continuous(limits = c(-1, 3.5)) +
    geom_point(data=observed, mapping=aes(y=Number, x=FA), inherit.aes = F) +
  geom_line(data=observed, mapping=aes(y=Number, x=FA), inherit.aes=F) +
    coord_flip() + 
    theme_ridges() + 
  geom_vline(xintercept=0, col="grey")

>>>>

---
title: 'Permanova: Real Data Analysis'
author: "Paul Harmon"
date: "6/26/2021"
output: html_document
---

@@

out <- MultiPermanova(CC2015_filter)
```

  #Visualize the results and store in a list
  sigplot2 <- ggplot(df3, aes(Holdout, PValues)) + geom_point(aes(color = Colors), size = 2) + geom_line(alpha = 0.5, color = "grey") + theme_bw() + ggtitle("CMD: P-Values From Adonis")
  sigplot2

  cc <- CC2015 %>% filter(BASIC2015 %in% c(15,16,17)) 
  cc$NAME[which(df3$PValues < 0.05)]

@@

## Scaled

@@

```{r, cache = TRUE
}
CC2015_filter_PC <- CC2015_filter %>% mutate(S.ER.DPC = S.ER.D/FACNUM, NONS.ER.DPC = NONS.ER.D/FACNUM, STAFFPC = PDNFRSTAFF/FACNUM)

@@

### 2021-08-21 Simulation Study Document V3.Rmd
Adding files from R

---
title: "Simulation Study Document"
author: "Paul Harmon"
date: "5/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

>>>>

---
title: "Simulation Study Document V4"
author: "Paul Harmon"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)

### 2021-08-21 Simulation Study Document.Rmd
Adding files from R

This doc contains simulation studies around the efficacy of a Permanova-based method for identifying influential points in a t-SNE or MDS map.  The document steps through a process of simulating data from 3 different groups (currently, these groups are clearly separated) with *3* influential points simulated on a standardized scale 5-6 standard deviations above the observed data.

The code below outlines the number of simulations run - currently this number is small for computational ease but will be increased (with the use of adaptive Permanova methods).

```{r}
n_sims = 100

#create a bunch of lists to store stuff later
plot_data_list <- list()
plot_tsne_results_list <- list()
plot_mds_results_list <- list()
plot_tsne_fstat_results_list <- list()
plot_mds_fstat_results_list <- list()
results_list <- list()
df3list = df2list = list()
```

The below chunk runs the simulations. Each iteration simulates a new (contaminated) dataset, runs the t-SNE and MDS mapping procedure, and tests the consistency of the configurations for each of the *n* holdout points. **This code is not run in this version.**

```{r original_sims, eval = FALSE, cache = TRUE, warning = FALSE, message = FALSE}

@@

  #### Simulate the Data

  g1 <- rmnorm(n = 10, mean = rep(20, 10), diag(25, 10)) %>%
    t() %>%
    data.frame()
  g2 <- rmnorm(n = 10, mean = rep(50, 20), diag(25, 20)) %>%
    t() %>%
    data.frame()
  g3 <- rmnorm(n = 10, mean = rep(80, 30), diag(25, 30)) %>%
    t() %>%
    data.frame()

@@

  df2 = tibble(Holdout = 1:length(x3$tsnelist), PValues = sapply(x3$tsnelist, pullPval), Fstat = sapply(x3$tsnelist, pullFstat))
  df2$Colors <- ifelse(df2$PValues < 0.05, TRUE, FALSE)
  df2list[[j]] = df2

  #### TSNE Plots########################################################
  #Visualize the results and store in a list
  sigplot <- ggplot(df2, aes(Holdout, PValues)) + geom_point(aes(color = Colors), size = 2) + geom_line(alpha = 0.5, color = "grey") + theme_bw() + ggtitle("t-SNE: P-Values From Adonis")
  plot_tsne_results_list[[j]] <- sigplot

  #Visualize the f stats and store in a list
  sigplotf <- ggplot(df2, aes(Holdout, Fstat)) + geom_point(aes(color = Colors), size = 2) + geom_line(alpha = 0.5, color = "grey") + theme_bw() + ggtitle("t-SNE: F Stats From Adonis")
  plot_tsne_fstat_results_list[[j]] <- sigplotf

  ### CMD Plots #########################################################
  df3 = tibble(Holdout = 1:length(x3$cmdlist), PValues = sapply(x3$cmdlist, pullPval), Fstat = sapply(x3$cmdlist, pullFstat))
    df3$Colors <- ifelse(df3$PValues < 0.05, TRUE, FALSE)
  df3list[[j]] = df3

  #Visualize the results and store in a list
  sigplot2 <- ggplot(df3, aes(Holdout, PValues)) + geom_point(aes(color = Colors), size = 2) + geom_line(alpha = 0.5, color = "grey") + theme_bw() + ggtitle("CMD: P-Values From Adonis")
  plot_mds_results_list[[j]] <- sigplot2

  #Visualize the f stats and store in a list
  sigplot2f <- ggplot(df3, aes(Holdout, Fstat)) + geom_point(aes(color = Colors), size = 2) + geom_line(alpha = 0.5, color = "grey") + theme_bw() + ggtitle("CMD: F Stats From Adonis")
  plot_mds_fstat_results_list[[j]] <- sigplot2f

@@

This chunk runs the simulation with slighlty increased variability in the three groups, and a restructuring of the influential points. Rather than simulating as outliers all on the positive side, I simulate one that has a large positive set of outlying points, one with a large negative set of outlying points, and one that simply 'bounces around' the multivariate space and breaks the grouping structure.

```{r sims_v2, eval = TRUE, cache = TRUE, warning = FALSE, message = FALSE}

  #### Simulate the Data with outliers adn a single group (simulates all from one group)

  g1 <- rmnorm(n = 10, mean = rep(20, 10), diag(45, 10)) %>%
    t() %>%
    data.frame()
  g2 <- rmnorm(n = 10, mean = rep(50, 20), diag(45, 20)) %>%
    t() %>%
    data.frame()
  g3 <- rmnorm(n = 10, mean = rep(80, 30), diag(45, 30)) %>%
    t() %>%
    data.frame()

  ## adds outliers - but not as a single group. Here we have an outlier that includes a really big one, a really small observation, and a very abnormal one
  add_outliers <- rbind(rnorm(10, 3, .5), rnorm(10, -3, .5), rnorm(10, c(rep(-2,3),rep(0,3),rep(2,3),0), .5)) %>% as_tibble()
  names(add_outliers) <- names(yes_structure_scale)

  #Visualize the data and store in a list
  datplot <- yes_structure_final %>% mutate(Id = 1:nrow(yes_structure_final), Alpha = ifelse(Group %in% 'Outlier', 1,0.8)) %>%  pivot_longer(1:10) %>% ggplot(aes(name, value, color = Group, group = Id, alpha = Alpha)) + geom_line() + geom_point() + ggtitle("Simulated Data with Outliers Added In") + guides(alpha = FALSE)
  plot_data_list[[j]] <- datplot

  df2 = tibble(Holdout = 1:length(x3$tsnelist), PValues = sapply(x3$tsnelist, pullPval), Fstat = sapply(x3$tsnelist, pullFstat))
  df2$Colors <- ifelse(df2$PValues < 0.05, TRUE, FALSE)
  df2list[[j]] = df2

  #### TSNE Plots########################################################
  #Visualize the results and store in a list
  sigplot <- ggplot(df2, aes(Holdout, PValues)) + geom_point(aes(color = Colors), size = 2) + geom_line(alpha = 0.5, color = "grey") + theme_bw() + ggtitle("t-SNE: P-Values From Adonis")
  plot_tsne_results_list[[j]] <- sigplot

  #Visualize the f stats and store in a list
  sigplotf <- ggplot(df2, aes(Holdout, Fstat)) + geom_point(aes(color = Colors), size = 2) + geom_line(alpha = 0.5, color = "grey") + theme_bw() + ggtitle("t-SNE: F Stats From Adonis")
  plot_tsne_fstat_results_list[[j]] <- sigplotf

  ### CMD Plots #########################################################
  df3 = tibble(Holdout = 1:length(x3$cmdlist), PValues = sapply(x3$cmdlist, pullPval), Fstat = sapply(x3$cmdlist, pullFstat))
    df3$Colors <- ifelse(df3$PValues < 0.05, TRUE, FALSE)
  df3list[[j]] = df3

  #Visualize the results and store in a list
  sigplot2 <- ggplot(df3, aes(Holdout, PValues)) + geom_point(aes(color = Colors), size = 2) + geom_line(alpha = 0.5, color = "grey") + theme_bw() + ggtitle("CMD: P-Values From Adonis")
  plot_mds_results_list[[j]] <- sigplot2

  #Visualize the f stats and store in a list
  sigplot2f <- ggplot(df3, aes(Holdout, Fstat)) + geom_point(aes(color = Colors), size = 2) + geom_line(alpha = 0.5, color = "grey") + theme_bw() + ggtitle("CMD: F Stats From Adonis")
  plot_mds_fstat_results_list[[j]] <- sigplot2f

# Results {.tabset}

The simulated datasets can be seen in the tabset below.

## PlotList
```{r}
library(ggpubr)
#ggarrange(plotlist = plot_data_list)
print(plot_data_list)
```

## Summary Metrics

We can start to look at some results below. The code generates the rate of correctly classified influential points for both the MDS (shown first) and t-SNE maps. The MDS correctly classifies the influential point *some* of the time, but the t-SNE method very rarely classifies the influential points as having an impact on the configuration.

@@

#print the head of the df lists
df3list %>% head(3)
### Identify the number of CORRECTLY Classified influential points
#extract the p-values from the last 3 from MDS
influential_correct <- function(temp){length(which(tail(temp, 3)$PValues < 0.05))/3}
mds_correct <- lapply(df3list, influential_correct)

# extract the pvalues from the last 3 of the t-SNE
tsne_correct <- lapply(df2list, influential_correct)

dat = tibble(MDS = unlist(mds_correct), TSNE = unlist(tsne_correct))
#plot(dat$MDS, dat$TSNE)

mds_correct %>% unlist() %>% mean() %>% kable(col.names = "MDS Percent Correctly Classified")
tsne_correct %>% unlist() %>% mean() %>% kable(col.names = "t-SNE Percent Correctly Classified")

Now we can look at the rate of misclassified non-influential points. These are the 60 'regular' points that are spuriously detected as outliers. Each percentage is calculated as a percentage of the 60 that are NOT small p-values (currenlty < 0.05 without a correction). We want a high percentage of correctly-identified non-influential points (although given the machinery of the hypothesis tests used, we would expect at least some spurious detections).

Both are relatively high percentages (as we would expect) - but the t-SNE hovers right around the expected 0.95 and MDS is higher.

### Identify rate of CORRECTLY Classified non-influential points
#extract the p-values from the last 3 from MDS
ni_correct <- function(temp){length(which(head(temp, 60)$PValues >= 0.05))/60}

correct_mds <- lapply(df3list, ni_correct)
#kable(correct_mds, col.names = "Correctly Identified Non-Influential: MDS")

# extract the pvalues from the last 3 of the t-SNE
correct_tsne <- lapply(df2list, ni_correct)
#kable(correct_tsne, col.names = "Correctly Identified Non-Influential: tSNE")

correct_mds %>% unlist() %>% mean() %>% kable(col.names = "MDS Percent Correctly Classified Non-Influential")
correct_tsne %>% unlist() %>% mean() %>% kable(col.names = "t-SNE Percent Correctly Classified Non-Influential")

@@

## MDS P Values

@@

## MDS F Stats

```{r}
plot_mds_fstat_results_list
```

@@

## t-SNE F Stats

```{r}
plot_tsne_fstat_results_list
```

>>>>

This doc contains simulation studies around the efficacy of a Permanova-based method for identifying influential points in a t-SNE or MDS map.

```{r}
n_sims = 10
```

```{r, cache = TRUE, warning = FALSE, message = FALSE}
plot_data_list <- list()
plot_tsne_results_list <- list()
plot_mds_results_list <- list()
results_list <- list()

@@

  #### Simulate the Data
  set.seed(34536)
  g1 <- rmnorm(n = 10, mean = rep(20, 10), diag(25, 10)) %>%
    t() %>%
    data.frame()
  g2 <- rmnorm(n = 10, mean = rep(50, 20), diag(25, 20)) %>%
    t() %>%
    data.frame()
  g3 <- rmnorm(n = 10, mean = rep(80, 30), diag(25, 30)) %>%
    t() %>%
    data.frame()

@@

  df2 = tibble(Holdout = 1:length(x3$tsnelist), PValues = sapply(x3$tsnelist, pullPval), Fstat = sapply(x3$tsnelist, pullFstat))
  df2$Colors <- ifelse(df2$PValues < 0.05, TRUE, FALSE)

  #Visualize the results and store in a list
  sigplot <- ggplot(df2, aes(Holdout, PValues)) + geom_point(aes(color = Colors), size = 2) + geom_line(alpha = 0.5, color = "grey") + theme_bw() + ggtitle("t-SNE: P-Values From Adonis")
  plot_tsne_results_list[[j]] <- sigplot

  df3 = tibble(Holdout = 1:length(x3$cmdlist), PValues = sapply(x3$cmdlist, pullPval), Fstat = sapply(x3$cmdlist, pullFstat))
  df3$Colors <- ifelse(df3$PValues < 0.05, TRUE, FALSE)

  #Visualize the results and store in a list
  sigplot2 <- ggplot(df2, aes(Holdout, PValues)) + geom_point(aes(color = Colors), size = 2) + geom_line(alpha = 0.5, color = "grey") + theme_bw() + ggtitle("t-SNE: P-Values From Adonis")
  plot_mds_results_list[[j]] <- sigplot2

@@

We can start to look at some results below. The code generates the rate of correctly classified influential points for both the MDS (shown first) and t-SNE maps.

```{r}
### Identify the number of CORRECTLY Classified influential points
#extract the p-values from the last 3 from MDS
length(which(tail(df3, 3)$PValues < 0.05))/3
# extract the pvalues from the last 3 of the t-SNE
length(which(tail(df2, 3)$Pvalues < 0.05))/3
```

Now we can look at the rate of misclassified non-influential points. These are the 60 'regular' points that are spuriously detected as outliers.

@@

### Identify rate of CORRECTLY Classified non-influential points
#extract the p-values from the last 3 from MDS
length(which(head(df3, 60)$PValues >= 0.05))/60 
# extract the pvalues from the last 3 of the t-SNE
length(which(tail(df2, 60)$PValues >= 0.05))/60
```

@@

## MDS

@@

@@

### 2021-08-21 Simulation-Study-Document-II.Rmd
Adding files from R

The below chunk runs the simulations. Each iteration simulates a new (contaminated) dataset, runs the t-SNE and MDS mapping procedure, and tests the consistency of the configurations for each of the *n* holdout points. **This code is not run in this version.**

@@

This chunk runs the simulation with slighlty increased variability in the three groups, and a restructuring of the influential points. Rather than simulating as outliers all on the positive side, I simulate one that has a large positive set of outlying points, one with a large negative set of outlying points, and one that simply 'bounces around' the multivariate space and breaks the grouping structure.

>>>>

The below chunk runs the simulations. Each iteration simulates a new (contaminated) dataset, runs the t-SNE and MDS mapping procedure, and tests the consistency of the configurations for each of the *n* holdout points.

@@

### 2021-08-21 Simulation-Study-Document.Rmd
Adding files from R

```{r}
n_sims = 10
```

@@

  df2 = tibble(Holdout = 1:length(x3$tsnelist), PValues = sapply(x3$tsnelist, pullPval), Fstat = sapply(x3$tsnelist, pullFstat))
  df2$Colors <- ifelse(df2$PValues < 0.05, TRUE, FALSE)

@@

  df3 = tibble(Holdout = 1:length(x3$cmdlist), PValues = sapply(x3$cmdlist, pullPval), Fstat = sapply(x3$cmdlist, pullFstat))
  df3$Colors <- ifelse(df3$PValues < 0.05, TRUE, FALSE)

>>>>

```{r}
n_sims = 100
```

@@

  df2 = tibble(Holdout = 1:length(x3$tsnelist), PValues = sapply(x3$tsnelist, pullPval))
  df2$Colors <- ifelse(df2$PValues < 0.05, TRUE, FALSE)

@@

  df3 = tibble(Holdout = 1:length(x3$cmdlist), PValues = sapply(x3$cmdlist, pullPval))
  df3$Colors <- ifelse(df3$PValues < 0.05, TRUE, FALSE)
